### 2014-05-21 writeup_and_notes/signif_statement.tex
edited signif down to 119

\documentclass{article}
\usepackage{graphicx}

\begin{document}
Meiosis is the time when alleles compete to ensure their transmission at the expense of organismal fitness, and many features of meiosis may have evolved to reduce competition. The observation that fertilization is often required for the completion of female meiosis appears incompatible with this view; sperm-acting alleles could distort female meiosis in favor, creating an additional conflict. Counterintuitively, we show that sperm are selected to ensure a fair female meiosis. This comes about because sperm immediately suffer the result of cheating, and so it is in sperm's interests to enforce fairness in female meiosis.  The requirement of fertilization to complete female meiosis may represent a long-term collaboration between sperm and mothers to ensure the production of fit offspring.
\end{document}