### 2014-05-21 writeup_and_notes/signif_statement.tex
yb [hopefully] improved  gcs sig statement while keeping it at 119 words

\documentclass{article}
\usepackage{graphicx}

\begin{document}
During female meiosis, alleles can take charge of their transmission, and those that bias meiosis in their favor can spread at a cost to organismal fitness. It is therefore believed that basic features of oogenesis likely evolved to control such renegade genes. The observation that female meiosis requires fertilization for its completion appears incompatible with this view; sperm-acting alleles could help like alleles distort  female meiosis in their favor, creating an additional conflict. Counterintuitively, we show that sperm succeed by enforcing fairness in females meiosis because this allows them to avoid pairing with low-fitness cheaters.  The requirement of fertilization to complete female meiosis may represent a long-term collaboration between sperm and mothers to ensure the production of fit offspring.
\end{document}